PoVaTiM: A Possibilistic Valid-Time Model \\ \vspace*{0.04truein}

### 2012-06-20 ijcis-pos-model/sources/temporalModel.tex
final draft

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated with an example. The information in the system is defined by the following elements:

\begin{definition}
\label{def:generalized-fuzzy-domain}
\textbf{Generalized fuzzy domain.}
$D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
\begin{equation}
\label{eq:generalized-fuzzy-domain}
D_G \subseteq \tilde \Pow\left(D \right)\cup \text{NULL}
\end{equation}
\end{definition}
The datatypes for $D_G$ are shown in table \ref{tbl:gefred-data-types}.

\begin{definition}
\label{def:typeof-domain}
\textbf{Typeof(D).}
The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$. For every datatype, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
\end{definition}

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:gefred-data-types}Data types}
%\centerline{\small TITLE}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{4cm} }\\
No. & Datatype \\ \hline
1 & A single scalar. \\
2 & A single number. \\
3 & A set of mutually exclusive possible scalar assignations. \\
4 & A set of mutually exclusive possible numeric assignations. \\
5 & A possibility distribution in a scalar domain. \\
6 & A possibility distribution in a numeric domain. \\
7 & A real number in $\left[0, 1 \right]$ referring to degree of matching. \\
8 & An \emph{UNKNOWN} value. \\
9 &  An \emph{UNDEFINED} value. \\
10 & A \emph{NULL} value. \\
\hline\\
\end{tabular}
}

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-domain}
\textbf{Generalized fuzzy temporal domain.}
If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
The Generalized Fuzzy Temporal Domain, $\T_G$ is
\begin{equation}
\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
\end{equation}
\end{definition}

@@

\begin{definition}
\textbf{Generalized fuzzy temporal relation.}
A generalized fuzzy temporal relation $R_{FTG}$ is given by:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

\begin{example}
\label{ex:sample-ft-database}
 Consider a library database as explained in example \ref{ex:library-database}. But now, the time is stored with imprecision. For simplicity, the ill-known time points are represented as triangular fuzzy numbers. An ill-known time point is given by $[dd/mm/yyyy,a,b]$. The values for $a$ and $b$ are integers. Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
\begin{align}
 \label{eq:head-ex}
\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
\nonumber
&  \left(\text{Title}:D_{\text{title}}\right),\\
\nonumber
& \left(\text{Loan}:D_{\text{Loan}}\right) \\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
\end{align}

@@

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample}Sample library database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c |p{0.35\columnwidth}}\\
\hline
& & & \centerline{PVP} \\
$\Head$ &  \textbf{ID}  & Title  &  Start Loan , End Loan\\
\hline
\multirow{3}{*}{$\Body$ } &  3 & Dracula & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
 &4 & Frankenstein & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
&5 & Harry Potter & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
\hline\\
\end{tabular}
}

\begin{definition}
\label{def:value-component}
\textbf{Value component.}
The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
\begin{align}
\label{eq:value-component}
R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
\nonumber
\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace 
\end{align}
\end{definition}

For example, in the case of the record with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Title}:D_{\text{title}}\right),\\
\nonumber
& \left(\text{Loan}:D_{\text{Loan}}\right), \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \left \lbrace 3:, \text{``Dracula``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
\end{align}

\begin{definition}
\label{def:compatibility-component}
\textbf{Compatibility component.}
The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
\begin{align}
\label{eq:compatibility-component}
R^{c}_{FTG} = \left \lbrace \Head^{c},\Body^{c} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{c} = \left \lbrace \left[C_{A_{G1}} \right], \ldots,  \left[C_{A_{Gn}} \right] \right \rbrace \\
\nonumber
\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace 
\end{align}
\end{definition}

@@

Analogously, it is possible to define the the value component for the temporal part and the compatibility component for the temporal part.

\begin{definition}
\label{def:generalized-primary-key}
\textbf{Generalized primary key.}
A generalized primary key, $K_G$ is a subset of the head:
\begin{align}
\label{eq:generalized-primary-key}
K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
\nonumber
K_G \subseteq \Head, K_G = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \right \rbrace
\end{align}
In this case, the function \emph{Typeof(ID) = 2} (see Definition \ref{def:typeof-domain} and Table \ref{tbl:gefred-data-types}). The primary key for the table is the attribute \emph{ID}, a unique number. Two different books have two different values for the \emph{ID} attribute.

In order to add valid-time support, the primary key should be re-defined. E.g., consider the library database. If the primary key is the \emph{ID} attribute, a book could not be loaned more than once. To resolve this problem, we extend the given primary key with a version identifier. 
%Thus, when a book is loaned, a new entry in the table is added. The \emph{ID} attribute identifies the physical book whereas

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\textbf{Generalized fuzzy temporal key.}
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
\nonumber
 \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\nonumber
\text{Typeof }\left(D_{ID} \right) = \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
 \nonumber
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \cup  \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
%\nonumber
 %\left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\end{align}

\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c | p{0.35\columnwidth} }\\
\hline
% & & & \multicolumn{2}{c}{PVP} \\
%& & & & \centerline{PVP}\\
{$\Head$}  &  &  & &(Start Loan\\
& \textbf{ID}  & \textbf{V} & Title& , End Loan)	\\
 \hline
\multirow{4}{*}{$\Body$}  & 3 & 001 & Dracula & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
& 4 & 001 & Frankenstein & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
& 5 & 001 & Harry Potter & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
 & 3 & 002 & Dracula & [4/4/2012, 3, 3] , UC \\
\hline\\
\end{tabular}
}

@@

The algebra for the data manipulation is defined in \cite{Medina1994}. The Generalized Fuzzy Relational Algebra manipulates relations like $R_{FG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify, Delete, } and \emph{Revise}. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.

It is important to notice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.  Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$). Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):

% \begin{itemize}
% \item 
\begin{definition}
\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time. In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
\begin{align}
\label{eq:stricly-consistent}
\forall I,J, \forall t \in \T :\\
%\nonumber
%\left(A, I \right) \in R \wedge \\
%\nonumber
%\left(A, J \right) \in R : \\
\nonumber
\pi_{I}(t) + \pi_{J}(t) \leq 1
\end{align}
\end{definition}

%\item 
\begin{definition}
 \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinguished among the following sub-types:
\end{definition}
% 	\begin{itemize}
% 	\item 
	      \begin{definition}
	      \textbf{Co-existence} Two versions of the same object $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:co-existence}
		      \forall I, \exists ! J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

	%\item 
	      \begin{definition}
	      \textbf{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:weak-consistence}
		      \forall I, \exists J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

@@

In the implementation of the DML operations we will consider a strictly consistent temporal database. This is a natural extension of the crisp temporal databases. This approach allows a easy implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.

Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.

@@

\begin{definition}
\label{def:pvp-current-in-relation}
\textbf{Current$\left(R, A \right)$.}
Consider an entity $A$ and a relation $R$ in a fuzzy temporal database. The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R$, that is it, $I = \left[X , \text{UNKNOWN} \right]$.

\begin{align}
\label{eq:pvp-current-in-relation}
\mbox{Current} \left(R, A \right) &=& \\ 
\begin{cases}
\nonumber
I & \mbox{ if } I = \left[X, UC \right] \\
%&  (A,I) \in R \wedge Y \in \text{UNKNOWN}\\
\emptyset & \mbox{ in any other case }
\end{cases}
\end{align}
\end{definition}

For example, consider the database in Table \ref{tbl:fuzzy-library-sample-valid-time}. The function Current$\left(R, A \right)$ with $A =\left(ID=3\right)$ returns time interval for the current version of $A$, $I = \left[ \left[4/4/2012, 3, 3 \right], UC \right]$. Nevertheless, the time interval for the entity with ID = 4 is the empty set.
Now it is possible to close the current version of an object by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.

\begin{definition}
\label{def:pvp-close-current-version}
\textbf{Close-current$\left(R, \left(A, J \right) \right)$.}
Consider an object $A$, the relation $R$ and a pvp, $J$. Then, the function close - current$\left(R, A, J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.

\begin{eqnarray}
\label{eq:pvp-close-current}
\text{Close-current} \left(R, \left(A, J \right) \right) =\\
\begin{cases}
\nonumber
\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
\nonumber
\mbox{ if } I = \mbox{ Current } \left(R, A \right) \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber R , \text{ in any other case}
\end{cases}
\end{eqnarray}
\end{definition}

>>>>

In this section we will formalize the model for possibilistic valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, the data manipulation language is defined.

\subsection{\label{subsec:temporal-model}The generalized temporal model}
The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model. This model is extended by adding valid-time support which will be illustrated through definitions and examples. The information in the system is defined by the following elements:

\begin{definition}
\label{def:generalized-fuzzy-domain}
\emph{Generalized fuzzy domain.}
Let $D$ be the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
\begin{equation}
\label{eq:generalized-fuzzy-domain}
D_G \subseteq \tilde \Pow\left(D \right)\cup \text{NULL}
\end{equation}
\end{definition}
The datatypes that can be used to represent $D_G$ are shown in table \ref{tbl:gefred-data-types}.

\begin{definition}
\label{def:typeof-domain}
\textbf{Typeof(A).}
Consider $D_G$ to be a generalized fuzzy domain and let $A \in D_G$ be a value for the domain. 
The function typeof$\left(A \right)$ returns the datatype associated with the domain $D_G$. For a given value $A$, the function returns a number in $\left[1, 10 \right]$ as shown in Table \ref{tbl:gefred-data-types}.
\end{definition}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:gefred-data-types}Data types}
%\centerline{\small TITLE}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{4cm} }\\
No. & Datatype \\ \hline
1 & A single scalar. \\
2 & A single number. \\
3 & A set of mutually exclusive possible scalar assignations. \\
4 & A set of mutually exclusive possible numeric assignations. \\
5 & A possibility distribution in a scalar domain. \\
6 & A possibility distribution in a numeric domain. \\
7 & A real number in $\left[0, 1 \right]$ referring to degree of matching. \\
8 & An \emph{UNKNOWN} value. \\
9 &  An \emph{UNDEFINED} value. \\
10 & A \emph{NULL} value. \\
\hline\\
\end{tabular}
} 
\end{samepage}

@@

\begin{definition}
\label{def:generalized-fuzzy-temporal-domain}
\emph{Generalized fuzzy temporal domain.}
Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
The Generalized Fuzzy Temporal Domain, $\T_G$ is
\begin{equation}
\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
\end{equation}
\end{definition}

@@

A generalized fuzzy relation is defined in \cite{Medina1994}. Here, we will extend the definition to a generalized fuzzy temporal relation.

\begin{definition}
\emph{Generalized fuzzy temporal relation.}
Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity. Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.  The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I. Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation. Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:

\begin{equation}
\label{eq:fuzzy-rel-def}
\left( R_{FTG}, \left(A , I\right) \right)
\end{equation}

A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
\label{def:generalized-fuzzy-temporal-relation}
\begin{equation}
\label{eq:generalized-fuzzy-temporal-relation}
R_{FTG} = \left(\Head, \Body \right)
\end{equation}
Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:

@@

\begin{example}
\label{ex:sample-ft-database}
 Consider a hospital database as explained in example \ref{ex:library-database}. But now, the time is stored with imprecision, because the date of arrival and departure are not precisely known. For simplicity, the ill-known time points are represented as triangular fuzzy numbers. An ill-known time point is given by $[dd/mm/yyyy,a,b]$. The values for $a$ and $b$ are integers. Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
\begin{align}
 \label{eq:head-ex}
\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
\nonumber
&  \left(\text{Name}:D_{\text{Name}}\right),\\
% \nonumber
% & \left(\text{Loan}:D_{\text{Loan}}\right) \\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
\end{align}

@@

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c |p{0.35\columnwidth}}\\
\hline
& & & \centerline{PVP} \\
$\Head$ &  \textbf{ID}  & Name  &  Arrival , Leave\\
\hline
\multirow{3}{*}{$\Body$ } &  3 & John & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
 &4 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
&5 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
\hline\\
\end{tabular}
}
\end{samepage}

\begin{definition}
\label{def:value-component}
\emph{Value component.}
The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
\begin{align}
\label{eq:value-component}
R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
\nonumber
\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace 
\end{align}
\end{definition}

For example, in the case of the record with ID = 3:
\begin{align}
%\label{eq:value-component-s}
\nonumber
\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Title}:D_{\text{title}}\right),\\
\nonumber
& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
\nonumber
\Body^ {v} &= \left \lbrace 3:, \text{``John``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
\end{align}

\begin{definition}
\label{def:compatibility-component}
\emph{Compatibility component.}
The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
\begin{align}
\label{eq:compatibility-component}
R^{c}_{FTG} = \left \lbrace \Head^{c},\Body^{c} \right \rbrace \\
\nonumber
\text{Where: } \\
\nonumber
\Head^{c} = \left \lbrace \left[C_{A_{G1}} \right], \ldots,  \left[C_{A_{Gn}} \right] \right \rbrace \\
\nonumber
\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace 
\end{align}
\end{definition}

@@

Analogously, it is possible to define both the name value component and the compatibility component for the temporal part.

%the value component for the temporal part and the compatibility component for the temporal part.

\begin{definition}
\label{def:generalized-primary-key}
\emph{Generalized primary key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$.
A generalized primary key, $K_G$ is a subset of the head:
\begin{align}
\label{eq:generalized-primary-key}
K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\end{align}
Subject to the following constraints:
\begin{align}
%\nonumber
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
\nonumber
K_G \subseteq \Head, K_G = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \right \rbrace
\end{align}
In this case, the function \emph{Typeof(ID) = 2} (see Definition \ref{def:typeof-domain} and Table \ref{tbl:gefred-data-types}). The primary key for the table is the attribute \emph{ID}, a unique number. Two different patients have two different values for the \emph{ID} attribute.

In order to add valid-time support, the primary key should be re-defined. E.g., consider the hospital database. If the primary key is the \emph{ID} attribute, a patient could not arrive to the hospital more than once. To resolve this problem, we extend the given primary key with a version identifier.

%a book could not be loaned more than once. 
%Thus, when a book is loaned, a new entry in the table is added. The \emph{ID} attribute identifies the physical book whereas

\begin{definition}
\label{def:generalized-fuzzy-temporal-key}
\emph{Generalized fuzzy temporal key.}
Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$. Let  $V$ be a new attribute called \emph{version}. 
A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head.
\begin{align}
\label{eq:generalized-fuzzy-temporal-key}
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
\nonumber
 \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\nonumber
\text{Typeof }\left(D_{ID} \right) = \\
\nonumber
s\in S \subseteq \left(1, \ldots, n \right) \\
\end{align}
Subject to the following constraints:
\begin{align}
\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
\nonumber
\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
\nonumber
\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
\end{align}
\end{definition}

@@

\begin{align}
 \nonumber
K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) , \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
%\nonumber
 %\left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
\end{align}

\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample hospital database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c | p{0.35\columnwidth} }\\
\hline
% & & & \multicolumn{2}{c}{PVP} \\
%& & & & \centerline{PVP}\\
{$\Head$}  &  &  & &(Arrival\\
& \textbf{ID}  & \textbf{V} & Name& ,Leave )	\\
 \hline
\multirow{4}{*}{$\Body$}  & 3 & 001 & John & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
& 4 & 001 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
& 5 & 001 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
 & 3 & 002 & John & [4/4/2012, 3, 3] , UC \\
\hline\\
\end{tabular}
}
\end{samepage}

@@

The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FTG}$. The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}. Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}. The operations implemented are: \emph{Insert, Modify, Delete, }. The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
%The algebra for the data manipulation is defined in .

It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is a value in the unit interval.  Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$). Thus, from the point of view of a fuzzy temporal database, there are the following choices (see Figure \ref{fig:tdb-class}):

% \begin{itemize}
% \item 
\begin{definition}
\textbf{Strictly consistent} It is not possible for two PVPs of the same entity to be valid at the same time. In other words, if $A$ is an entity, $R_{FTG}$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
\begin{align}
\label{eq:stricly-consistent}
\forall I,J, \forall t \in \T :\\
%\nonumber
%\left(A, I \right) \in R \wedge \\
%\nonumber
%\left(A, J \right) \in R : \\
\nonumber
\pi_{I}(t) + \pi_{J}(t) \leq 1
\end{align}
\end{definition}

%\item 
\begin{definition}
 \textbf{Ill-consistent} The overlapping of two PVPs of the same entity do overlap. It can be distinguished among the following sub-types:
\end{definition}
% 	\begin{itemize}
% 	\item 
	      \begin{definition}
	      \textbf{Co-existence} Two versions of the same entity $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:co-existence}
		      \forall I, \exists ! J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

	%\item 
	      \begin{definition}
	      \textbf{Weak-consistence} Several versions of the same entity $A$ may exist at the same time.
		      \begin{align}
		      \label{eq:weak-consistence}
		      \forall I, \exists J, \exists t \in \T : \\
		      \nonumber
		      \pi_{I}(t) + \pi_{J}(t) \geq 1
		      \end{align}
	      \end{definition}

@@

In the implementation of the DML operations we will consider a strictly consistent temporal database. This is a natural extension of the crisp temporal databases. This approach allows an implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.

Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} are the basis for the following auxiliary functions.

@@

\begin{definition}
\label{def:pvp-current-in-relation}
\textbf{Current$\left(R_{FTG}, A \right)$.}
Consider an entity $A$ and a relation $R_{FTG}$ in a fuzzy temporal database. The function Current$\left(R_{FTG}, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R_{FTG}$, that is, $I = \left[X , \text{UNKNOWN} \right]$.

\begin{align}
\label{eq:pvp-current-in-relation}
\mbox{Current} \left(R_{FTG}, A \right) &=& \\ 
\begin{cases}
\nonumber
I & \mbox{ if } I = \left[X, UC \right] \\
%&  (A,I) \in R \wedge Y \in \text{UNKNOWN}\\
\emptyset & \mbox{ in any other case }
\end{cases}
\end{align}
\end{definition}

For example, consider the database in Table \ref{tbl:fuzzy-library-sample-valid-time}. The function Current$\left(R_{FTG}, A \right)$ with $A =\left(ID=3\right)$ returns time interval for the current version of $A$, $I = \left[ \left[4/4/2012, 3, 3 \right], UC \right]$. Nevertheless, the time interval for the entity with ID = 4 is the empty set.
Now it is possible to close the current version of an entity by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.

\begin{definition}
\label{def:pvp-close-current-version}
\textbf{Close-current$\left(R_{FTG}, \left(A, J \right) \right)$.}
Consider an entity $A$, the relation $R_{FTG}$ and a pvp, $J$. Then, the function close - current$\left(R_{FTG}, A, J \right)$ closes any current version of the entity $A$ if it exists and add the new version $\left(A, J \right)$.

\begin{eqnarray}
\label{eq:pvp-close-current}
\text{Close-current} \left(R_{FTG}, \left(A, J \right) \right) =\\
\begin{cases}
\nonumber
\big \lbrace R_{FTG} - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
\nonumber
\mbox{ if Current } \left(R_{FTG}, A \right) \neq \emptyset \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber R , \text{ in any other case}
\end{cases}
\end{eqnarray}
\end{definition}
